Handling missing data in smartphone location logs

### 2017-12-01 thesisReport.Rmd
added aggregation section with plot, figure table references

@@

Google location history provides a measure of accuracy that is given in meters such that it represents the radius of a 67% confidence circle. In the example data set the distribution of accuracy is highly right skewed, with a median of 28, $\mu = 127$ and the maximum value at 26 km. @palmius_detecting_2017 note that in their Android based custom logs inaccurate location values are interspersed between more accurate location values at higher sample rates per hour. We observe similar patterns in secondary logs. Figure 1 shows how accuracy can vary as a function of user behaviour, time and location. Inaccurate measures are often followed by more accurate measures.  Most notably, low accuracy often (but not always) is  associated with movement (Figure 2). Stationary accuracy varies depending on phone battery level, wifi connection and user phone use. There are several recurring low-accuracy points, possibly the result of cell-phone tower triangulation.

@@

Over 54% of the data is missing for the entire duration of the log. This may be misleading as there are several long periods with no measurements whatsoever (see Figure X a). For days which were not entirely missing, approximately 22% of all five minute segments were missing. The structure of missingness of a day with measurements is shown in Figure X b. As you can see, there are several long periods over the course of the log for which there are no measurements. Moreover, even during a single day there are continuous periods where there is missing data, mostly during the late hours of the night in this case.

>>>>

### Aggregate measures

Social scientists are most interested in aggregating spatiotemporal data to more socially relevant information, such as distance travelled. As we discussed earlier, aggregations without data processing can be highly biased. However, as an example we calculate time spent at home of the user for each day of the week in the month of February (\@ref(fig:aggrePlot)). Between individuals, time spent at home has been found to be a reliable predictor of extraversion [@harari_using_2016].

```{r, aggrePlot,echo = FALSE,fig.cap="Proportion of time spent at home in February 2017. We estimate this by calculating the mean lattitude and longitude for every 5 minute time period in the month. Missing values are filled in by the previous observed value. Behavioural trends are evident, the user spends more time at home on weekends than on weekdays. Moreover, the variance is greater on weekends, due to travel.",out.width = '100%'}
include_graphics("img/timeUse.png")
```

@@

Google location history provides a measure of accuracy that is given in meters such that it represents the radius of a 67% confidence circle. In the example data set the distribution of accuracy is highly right skewed, with a median of 28, $\mu = 127$ and the maximum value at 26 km. @palmius_detecting_2017 note that in their Android based custom logs inaccurate location values are interspersed between more accurate location values at higher sample rates per hour. We observe similar patterns in secondary logs. \@ref(fig:accuracyPlot) shows how accuracy can vary as a function of user behaviour, time and location. Inaccurate measures are often followed by more accurate measures.  Most notably, low accuracy often (but not always) is  associated with movement (\@ref(fig:accuracyPlot2)). Stationary accuracy varies depending on phone battery level, wifi connection and user phone use. There are several recurring low-accuracy points, possibly the result of cell-phone tower triangulation.

@@

Over 54% of the data is missing for the entire duration of the log. This may be misleading as there are several long periods with no measurements whatsoever (see Figure \@ref(fig:longMeasurementsPerDay)). For days which were not entirely missing, approximately 22% of all five minute segments were missing. The structure of missingness of a day with measurements is shown in Figure \@ref(fig:measurementsPerDay). As you can see, there are several long periods over the course of the log for which there are no measurements. Moreover, even during a single day there are continuous periods where there is missing data, mostly during the late hours of the night in this case.
